---
title: "Live Session 5 Assigment: Questions 1, 2, and 3"
author: "Manuel Rosales"
date: "February 8, 2018" 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1: Data Munging

a) First, import the .txt file into R so you can process it. Assign the resulting data frame to an object, df, that consists of three columns with human-readable column names for each

```{r Question1a}
# Question 1a
setwd("~/Library/Mobile Documents/com~apple~CloudDocs/DataScience@SMU/MSDS_DDS/DoingDDSGitHub/Unit_05")
df <- read.table("yob2016.txt", header = FALSE, sep = ";", col.names = c("FirstName","Gender","Count"))
names(df) # Confirm column names
head(df,10) # Review head of df
```

b) Display the summary and structure of df

```{r Question1b}
# Question 1b
summary(df)
str(df)
```

c) One name was entered twice and misspelled. He accidentally put three y's at the end of the name.  Write an R command to figure out which name it is and display it.

d) Upon finding the misspelled name, please remove this particular observation, as the client says it's redundant.  Save the remaining dataset as an object: y2016 




